Type Checker Generation using Reference Attribute Grammars

### 2023-11-06 MSc.tex
Fully rewrite implementation section

The type checker generation is based on a single-responsibility principle, where each node of the abstract syntax tree is responsible for generating the code it corresponds to.
Once typing rules have been parsed, we end up with an abstract syntax tree, in the form of figure \ref{trastspec}, where the \verb|RuleSet| is the root node.
The \verb|RuleSet| node consists of a list of \verb|Rule| nodes, and begins by sorting each rule by which object language node it describes.

The \verb|RuleSet| node then begins by generating the outline of a JastAdd aspect \verb|TypingRules|, and the definition for the method \verb|.type()| for each object language node.
It then calls on the \verb|Rule| node to fill in the implementation of the method.

A \verb|Rule| consists of a name, a conclusion, and a possibly empty list of premises that need to be true for the conclusion to be returned.
In the simplest case of an empty list of premises, the entire content of the rule is generated by the conclusion.
For rules containing premises, the code becomes more involved.
Firstly, any variables or type variables used in the rule need to be declared, followed by a check to ensure each use of a type variable represents a node with the same type.
Then it generates an if-statement corresponding to our premises, by letting each \verb|Premise| generate a boolean expression, all of which get joined by the AND operator.

Both \verb|Premise|s and \verb|Conclusion|s are represented by the same abstract class \verb|Formula|, which currently contains a single implementation \verb|HasType|, consisting of an expression and a type.
Depending on whether it is used as a premise or conclusion, it can either represent that the expression must have the specified type or that the expression should be given the specified type.

\subsection{Type checker design}
The provided typing rules compile into \verb|type()| methods, which return a value from the Type interface.
For the most simple typing rules, this method's implementation will consist of a single return statement.

Types are declared in an AST specification, currently provided by the template, allowing simple type hierarchies.

>>>>

The type checker generation is based on a single-responsibility principle, where each node of the abstract syntax tree is responsible for generating the code it corresponds to.
Once typing rules have been parsed, we end up with an abstract syntax tree, in the form of figure \ref{trastspec}, where the \verb|RuleSet| is the root node.

\subsubsection{Aspect generation}
The \verb|RuleSet| node is responsible for generating the overall structure of the output file, a JastAdd aspect called \verb|TypingRules|.
It sorts its \verb|Rule| children by which nodes they apply to, so that if multiple tying rules type the same node, they can be inserted into the same method implementation.

\subsubsection{Method implementation}
The implementation of a typing rule can range from a single return statement to a system of declarations of variables, type variables and control flow.

\subsubsection{Parameters}
When the typing rules concern a node with parameters, these can be named arbitrarily, and a variable with the same name will be instantiated at the beginning of the method, fetching the corresponding child node by index.
These variables are then utilised in later premises or type variable declarations.

\subsubsection{Premises/control flow}
The rule's premises are translated into an if-statement, which checks all types using a method \verb|matches| added to all Type nodes.
Its current implementation merely checks that the Type nodes are part of the same class, though it would be possible to override this with deeper comparisons for specific types.

\subsubsection{Type variables}
In the case of type variables, the generated code takes a dynamic approach to simplify the implementation.
Each type variable that appears in the conclusion is instantiated to a null value, and the code generated by each premise first checks if the value is still null and either assigns the variable its own type or tests to make sure the existing type matches its own.

\subsubsection{Error handling}
The current error handling consists of throwing a runtime exception at the first discovered type error, ceasing the type checking immediately.
This method was chosen for its simplicity of implementation, rather
